Model-based pretreatment for regression with correlated variables.

### 2014-04-09 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@518 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

{\bf Abstract.} Linear regression outcomes are known to be damaged by highly correlated covariates. However many modern datasets are expected to convey more and more highly correlated covariates due to the global increase of the amount of variables in datasets. We propose to explicitly model the correlations by a family of linear regressions between the covariates, some covariates explaining others. It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model, easily understandable and from which it is possible to perform standard linear estimation methods including variables selection procedures for instance. The structure of correlations is found with an MCMC algorithm. An R package (\textsc{CorReg}) available on the CRAN implements this new method which will be illustrated on both simulated datasets and real-life datasets from steel industry where correlated variables are frequent.
\smallskip

@@

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made, from industry (ballistic models to calibrate the process) to sociology (predicting some numerical properties of a population).
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators, even without bias from modelisation. 
In this paper, vectors and matrices are in bold letters to distinguish them from scalars. We note the linear regression model:
\begin{equation}
		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}
	\end{equation}
	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated with Ordinary Least Squares (\textsc{OLS}). %As shown in section \ref{sectionOLS}, 
	Estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
Variance of $\boldsymbol{\hat\beta}$ increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model: the more covariates you have the greater variance you get.
	\item The correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}

>>>>

{\bf Abstract.} Linear regression outcomes are known to be damaged by highly correlated covariates. However many modern datasets are expected to convey more and more highly correlated covariates due to the global increase of the amount of variables in datasets. We propose to explicitly model the correlations by a family of linear regressions between the covariates, some covariates explaining others. It allows then to obtain by marginalisation on the explained covariates a parsimonious correlation-free regression model, easily understandable. This pretreatment by variable selection gives then the possibility to perform standard linear estimation methods including variables selection procedures for instance. The structure of correlations is found with an MCMC algorithm. An R package (\textsc{CorReg}) available on the CRAN implements this new method which will be illustrated on both simulated datasets and real-life datasets from steel industry where correlated variables are frequent.
\smallskip

@@

When one wants to explain a phenomenon based on some covariates, the first statistical method tried frequently is the linear regression. It provides a predictive model with a good interpretability even  for non-statistician and is simple to learn. Therefore, linear regression is used in nearly all the fields where statistics are made \cite{montgomery2012introduction}, from astronomy \cite{isobe1990linear} to sociology \cite{longford2012revision}.
Linear regression is a very classic situation but sometimes it has to face an also classical problem: the variance of the estimators. We note the linear regression model:
\begin{equation}
		\boldsymbol{Y}_{|\boldsymbol{X}}=\boldsymbol{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon}
	\end{equation}
	where $\boldsymbol{X}$ is the $n\times p$ matrix of the explicative variables, $\boldsymbol{Y}$ the  $n\times 1$ response vector and $\boldsymbol{\varepsilon} \sim \mathcal{N}(\boldsymbol{0},\sigma_Y^2\boldsymbol{I_n})$ the noise of the regression, with $\boldsymbol{I_n}$ the $n$-sized identity matrix and $\sigma_Y \in \mathcal{R}$. The $p\times 1$ vector $\boldsymbol{\beta}$ is the vector of the coefficients of the regression, that can be estimated with Ordinary Least Squares (\textsc{OLS}). %As shown in section \ref{sectionOLS}, 
	Estimation of the parameters requires the inversion of a matrix which will be ill-conditioned or even singular if some covariates depend linearly from each other. 
Variance of $\boldsymbol{\hat\beta}$ increases based on two aspects :
\begin{itemize}
	\item The dimension $p$ (number of covariates) of the model: the more covariates you have the greater variance you get.
	\item The correlations within the covariates: strongly correlated covariates give bad-conditioning and increase variance of the estimators .
\end{itemize}
